---
title: "ISLR"
output:
  html_document:
    df_print: paged
autor: René Brinkhuis
---

# EXERCISE 8
This exercise relates to the College data set, which can be found in the file `college.csv`. It contains a number of variables for 777 different universities and colleges in the US. The variables are:

- `Private` : Public/private indicator
- `Apps` : Number of applications received
- `Accept` : Number of applicants accepted
- `Enroll` : Number of new students enrolled
- `Top10perc` : New students from top 10 % of high school class
- `Top25perc` : New students from top 25 % of high school class
- `F.Undergrad` : Number of full-time undergraduates
- `P.Undergrad` : Number of part-time undergraduates
- `Outstate` : Out-of-state tuition
- `Room.Board` : Room and board costs
- `Books` : Estimated book costs
- `Personal` : Estimated personal spending
- `PhD` : Percent of faculty with Ph.D.’s
- `Terminal` : Percent of faculty with terminal degree
- `S.F.Ratio` : Student/faculty ratio
- `perc.alumni` : Percent of alumni who donate
- `Expend` : Instructional expenditure per student
- `Grad.Rate` : Graduation rate

Before reading the data into `R` , it can be viewed in Excel or a text editor.

## 8a
Use the `read.csv()` function to read the data into `R`. Call the loaded data `college` . Make sure that you have the directory set to the correct location for the data.

```{r}
College <- read.csv('data/college.csv')
```

Or use the data from the `ISLR` package.

```{r, message=FALSE}
require(ISLR)
data(College)
```

```{r}
str(College)
```

## 8b
Look at the data using the `fix()` function. You should notice that the first column is just the name of each university. We don’t really want `R` to treat this as data. However, it may be handy to have these names for later. Try the following commands:

> rownames(College) = College[, 1]  
> fix(College)  

```{r}
# preproces data
College <- cbind(rownames(College), data.frame(College, row.names=NULL))
colnames(College)[1] <- "College"
fix(College)
```

```{r}
# convert column to index
rownames(College) <- College[, 1]
College <- College[, -1]
fix(College)
```

Now you should see that the first data column is `Private`. Note that another column labeled `row.names` now appears before the `Private` column. However, this is not a data column but rather the name that `R` is giving to each row.

## 8c [i]
Use the `summary()` function to produce a numerical summary of the variables in the data set.

```{r}
summary(College)
```

## 8c [ii]
Use the `pairs()` function to produce a scatterplot matrix of the first ten columns or variables of the data. Recall that you can reference the first ten columns of a matrix `A` using `A[,1:10]`.

```{r}
pairs(College[, 1:10])
```

## 8c [iii]
Use the `plot()` function to produce side-by-side boxplots of `Outstate` versus `Private`.

```{r}
boxplot(Outstate~Private, data=College, xlab="Private", ylab="Outstate")
```

## 8c [iv]
Create a new qualitative variable, called `Elite`, by binning the `Top10perc` variable. We are going to divide universities into two groups based on whether or not the proportion of students coming from the top 10% of their high school classes exceeds 50%.

> Elite <- rep("No", nrow(College))  
> Elite[College$Top10perc>50] <- "Yes"  
> Elite <- as.factor(Elite)  
> College <- data.frame(College, Elite)  

```{r}
Elite <- rep("No", nrow(College))
Elite[College$Top10perc>50] <- "Yes"
Elite <- as.factor(Elite)
College <- data.frame(College, Elite)
```

Use the `summary()` function to see how many elite universities there are.

```{r}
summary(College)
```

Now use the `plot()` function to produce side-by-side boxplots of `Outstate` versus `Elite`.

```{r}
boxplot(Outstate~Elite, data=College, xlab="Elite", ylab="Outstate")
```

## 8c [v]
Use the `hist()` function to produce some histograms with differing numbers of bins for a few of the quantitative variables. You may find the command `par(mfrow=c(2, 2))` useful: it will divide the print window into four regions so that four plots can be made simultaneously. Modifying the arguments to this function will divide the screen in other ways.

```{r}
par(mfrow=c(2, 2))
hist(College$Apps, breaks=10, main="Apps")
hist(College$Enroll, breaks=20, main="Enroll")
hist(College$Expend, breaks=30, main="Expend")
hist(College$Outstate, breaks=40, main="Outstate")
```

# EXERCISE 9
This exercise involves the `Auto` data set studied in the lab. Make sure that the missing values have been removed from the data.

## 9b
What is the range of each quantitative predictor? You can answer this using the `range()` function.

```{r}
range(Auto$mpg)
range(Auto$cylinders)
range(Auto$displacement)
range(Auto$horsepower)
range(Auto$weight)
range(Auto$acceleration)
range(Auto$year)
```

## 9c
What is the mean and standard deviation of each quantitative predictor?

```{r}
sapply(Auto[,1:7], mean)
sapply(Auto[,1:7], sd)
```

## 9d
Now remove the 10th through 85th observations. What is the range, mean, and standard deviation of each predictor in the subset of the data that remains?

```{r}
Auto1085 <- Auto[, -(8:9)] # quantitative predictors only
Auto1085 <- Auto1085[-(10:85), ] # remove rows
sapply(Auto1085, range)
sapply(Auto1085, mean)
sapply(Auto1085, sd)
```

## 9e
Using the full data set, investigate the predictors graphically, using scatterplots or other tools of your choice. Create some plots highlighting the relationships among the predictors. Comment on your findings.

```{r}
pairs(Auto[,1:7])
```

## 9f
Suppose that we wish to predict gas mileage (`mpg`) on the basis of the other variables. Do your plots suggest that any of the other variables might be useful in predicting `mpg`? Justify your answer.

```{r}
cor(Auto$mpg, Auto$cylinders)
cor(Auto$mpg, Auto$displacement)
cor(Auto$mpg, Auto$horsepower)
cor(Auto$mpg, Auto$weight)
cor(Auto$mpg, Auto$acceleration)
cor(Auto$mpg, Auto$year)
```

# EXERCISE 10
To begin, load in the `Boston` data set. The Boston data set is part of the `MASS` library in `R`.  

Read about the data set: `?Boston`  

How many rows are in this data set? How many columns? What do the rows and columns represent?

## 10a
```{r, message=FALSE}
require(ISLR)
require(MASS)
data(Boston)
str(Boston)
```

```{r}
?Boston
```

```{r}
nrow(Boston)
ncol(Boston)
```

## 10b
Make some pairwise scatterplots of the predictors (columns) in this data set. Describe your findings.

```{r}
pairs(Boston)
```

## 10c
Are any of the predictors associated with per capita crime rate? If so, explain the relationship.

```{r, message=FALSE}
require(ggplot2)
require(reshape2)

bosmelt <- melt(Boston, id="crim")
ggplot(bosmelt, aes(x=value, y=crim)) +
  facet_wrap(~variable, scales="free") + 
  geom_point()
```

```{r}
corrmatrix <- cor(Boston, use="complete.obs")[1,]
corrmatrix[corrmatrix > 0.5 | corrmatrix < -0.5][-1]
```

## 10d
Do any of the suburbs of Boston appear to have particularly high crime rates? Tax rates? Pupil-teacher ratios? Comment on the range of each predictor.

```{r, message=FALSE}
require(ggplot2)
n = nrow(Boston)
```

```{r}
ggplot(Boston, aes(x=1:n, y=crim)) + geom_point()
```

```{r}
ggplot(Boston, aes(x=1:n, y=tax)) + geom_point()
```

```{r}
ggplot(Boston, aes(x=1:n, y=ptratio)) + geom_point()
```

## 10e
How many of the suburbs in this data set bound the Charles river?

```{r}
table(Boston$chas)
```

## 10f
What is the median pupil-teacher ratio among the towns in this data set?

```{r}
median(Boston$ptratio)
```

## 10g
Which suburb of Boston has lowest median value of owner-occupied homes? What are the values of the other predictors for that suburb, and how do those values compare to the overall ranges for those predictors? Comment on your findings.

```{r}
Boston[Boston$medv == min(Boston$medv),]
```

```{r}
sapply(Boston, quantile)
```

## 10h
In this data set, how many of the suburbs average more than seven rooms per dwelling? More than eight rooms per dwelling? Comment on the suburbs that average more than eight rooms per dwelling.

```{r}
nrow(Boston[Boston$rm > 7,])
```

```{r}
nrow(Boston[Boston$rm > 8,])
```
